Temporality is Important for Networks of Hospital Patient Transfers

### 2024-11-14 reports/report.qmd
Add diagram

---
title: "Temporality is Important for Networks of Hospital Patient Transfers"
authors:
    - name: David Wu
    - name: Michael Lydeamore
    - name: Andrew Stewardson
    - name: Tjibbe Donker
    - name: Nic Geard
    - name: "******"
date: last-modified
format:
    html: default
    pdf:
        include-in-header:
            - text: |
                \usepackage{mathtools}
        papersize: a4
bibliography: refs.bib
csl: plos.csl
jupyter: python3
execute:
    echo: false
    warning: false
    error: false
---

@@

Victoria, Australia, has a population of 7.0 million people [@ABS_state_populations]. Its hospital system consists of more than 300 hospital campuses, which vary in size, case mix and services.

Hospital admissions data was sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official].
In particular, admissions data from [] to [] ([] days) was used in this study. All Victorian hospital admissions during this study period are included. The total number of admissions is []
This data was aggregated into time-discretised temporal networks (described below), and exported from the database; further analysis as presented in this paper is performed on external machines with no access to identifiable protected health information.

@@

## Network Modelling {#sec-network-modelling}

@@

## Explicitly Temporal Models
Previous analyses of movement and spread of contagions through these networks have made the assumption that the weight $w_{ij}$ for a given edge does not change over time, i.e. the network is static.
This implies that the rate movement of patients between hospitals is Poisson, or that the time between successive admissions for a given patient is exponentially distributed. Empirically, this is poor modelling choice, since the readmission window, as shown in @fig-readmission-window, is definitely not exponentially distributed, and has a significant heavy tail. A naïve static approximation that matches the mean readmission window will be inaccurate; in the following, we will attempt to resolve this problem in three increasingly granular ways:

@@

<- Description of temporal networks and models>

<- Description of temporally reductive networks and models>

@@

### Networks and Projections

Patient transfers can be considered as edges from some source hospital $x_i$ at some time $t_i$ to a target hospital $x_j$ at some time $t_j$. 
Because we cannot export exact times, we instead discretise time.
This is done by treating movements involving some arbitrary hospital $x_i$ between times $t_i$ and $t_i + W$ as instead involving a single node at $v_i \coloneqq (x_i, t_i)$. 
Then we can have a temporal network that can be treated as a static network as in @sec-network-modelling, i.e. we have edge weights $w_{ij}$ along the edge $e_{ij} \coloneqq (v_i, v_j)$, but with some explicit temporal representation.

@@

@@

# Appendix

>>>>

---
title: "Temporality is Important for Networks of Hospital Patient Transfers"
authors:
    - name: David Wu
    - name: Michael Lydeamore
    - name: Andrew Stewardson
    - name: Tjibbe Donker
    - name: Nic Geard
    - name: "******"
date: last-modified
format:
    html: default
    pdf:
        include-in-header:
            - text: |
                \usepackage{mathtools}
        papersize: a4
bibliography: refs.bib
csl: plos.csl
filters:
    - diagram
jupyter: python3
execute:
    echo: false
    warning: false
    error: false
---

@@

Victoria, Australia, has a population of 7.0 million people [@ABS_state_populations]. Its hospital system consists of $N_x = 338$ hospital campuses, which vary in size, case mix and services.

Hospital admissions data was sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official].
In particular, admissions data from 1 January 2011 to 9 December 2020 (3631 days) was used in this study. All Victorian hospital admissions during this study period are included. The total number of admission records is 26876787.
This data was aggregated into time-discretised temporal networks (described below), and exported from the database; further analysis as presented in this paper is performed on external machines with no access to identifiable protected health information.

@@

## Existing Network Modelling {#sec-network-modelling}

@@

## Explicitly Temporal Models
Previous analyses of movement and spread of contagions through these networks have made assumptions that result in the weight $w_{ij}$ of any given edge being constant over time, i.e. the network is static. [.....Typically, this is a result of performing an aggregation, and approximating the patient movement between hospitals as being Poisson, or equivalently that the residence time  hospitals being exponentially distributed......]
As shown in @fig-readmission-window, these approximations lead to misleading dynamics on the hospital system of Victoria: the empirical distribution of inter-transfer times [....]. A naïve static approximation that matches the mean readmission window will be inaccurate; in the following, we will attempt to resolve this problem in three increasingly granular ways:

@@

@@

Patient transfers can be considered as edges from some source hospital $x_i$ at some time $t_i$ to a target hospital $x_j$ at some time $t_j$. 
Because we cannot export exact times, we instead discretise time.
This is done by treating movements involving some arbitrary hospital $x_i$ between times $t_i$ and $t_i + W$ as instead involving a single node at $v_i \coloneqq (x_i, t_i)$.

<!-- ::: {#fig-}
```{.tikz}
\begin{tikzpicture}
\tikzstyle{every node}=[font=\LARGE]
\draw [short] (2.75,14.75) -- (16.5,14.75);
\draw [short] (2.75,11) -- (16.5,11);
\draw [dashed] (3.5,16.25) -- (3.5,9.25);
\draw [dashed] (8.25,16.5) -- (8.25,9.25);
\draw [dashed] (12.25,16.5) -- (12.25,9.25);
\draw [dashed] (16,16.5) -- (16,9.25);
\draw [->, >=Stealth] (7.25,14.75) -- (9.25,11);
\draw [->, >=Stealth] (8.75,14.75) -- (13.25,11);
\draw [->, >=Stealth] (9.5,11) -- (10.75,14.75);
\draw [->, >=Stealth] (4.25,11) -- (5,14.75);
\draw [->, >=Stealth] (5.5,14.75) -- (5.5,11);
\draw [->, >=Stealth] (6.25,11) -- (7,14.75);
\draw [->, >=Stealth] (10,14.75) .. controls (11.75,15.75) and (12.5,15.75) .. (14.25,14.75) ;
\draw [->, >=Stealth] (15,11) -- (15,14.75);
\end{tikzpicture}
``` -->

<!-- ::: {#fig-time-discretisation}

```{.tikz}
%%| caption="This is an image, created by **TikZ i.e. LaTeX**."
%%| additionalPackages="\usepackage{adjustbox}"
\usetikzlibrary{arrows}

\resizebox{16cm}{!}{%
\begin{tikzpicture}[auto]
    \tikzstyle{every node}=[font=\LARGE]
    \draw [short] (0.275,1.475) -- (1.65,1.475);
    \draw [short] (0.275,1.1) -- (1.65,1.1);
    \draw [dashed] (0.35,1.65) -- (0.35,0.925);
    \draw [dashed] (0.825,1.65) -- (0.825,0.925);
    \draw [dashed] (1.225,1.65) -- (1.225,0.925);
    \draw [dashed] (1.6,1.65) -- (1.6,0.925);
    \draw [O-Latex] (0.725,1.475) -- (0.925,1.1);
    \draw [->, >=Stealth] (0.875,1.475) -- (1.325,1.1);
    \draw [->, >=Stealth] (0.95,1.1) -- (1.075,1.475);
    \draw [->, >=Stealth] (0.425,1.1) -- (0.5,1.475);
    \draw [->, >=Stealth] (0.55,1.475) -- (0.55,1.1);
    \draw [->, >=Stealth] (0.625,1.1) -- (0.7,1.475);
    \draw [->, >=Stealth] (1.0,1.475) .. controls (1.175,1.575) and (1.25,1.575) .. (1.425,1.475) ;
    \draw [->, >=Stealth] (1.5,1.1) -- (1.5,1.475);
\end{tikzpicture}
}
```

Example schematic for time discretisation
::: -->

Then we can have a temporal network that can be treated as a static network as in @sec-network-modelling, i.e. we have edge weights $w_{ij}$ along the edge $e_{ij} \coloneqq (v_i, v_j)$, but with some explicit temporal representation.

@@

![An example of a small 2-layer static network](diagrams/precompiled/BiLayer_ManimCE_v0.18.1.png){#fig-bilayer}

@@

{{< pagebreak >}}
# Appendix
